\chapter{Introduction}

\newpar The idea of creating mass-produced autonomous vehicle has seen increased interest from researchers and car manufacturers in the last decade.

The idea of inventing an autonomous, self-driving car is not a new thing. Dreaming up future societies where people would focus on other things while their cars drove them from A to B, has been a trope of science-fiction for decades\footnote{The popular TV show, \textit{The Jetsons} is one of the more well-known examples of having autonomous vehicles. \cite{wiki:jetsons}}. During the last decade these ideas are finally being realised with research projects such as Google's ongoing \textit{Self Driving Car Project}, previous research projects such as the entries of Carnegie Melon and Stanford in the \textit{DARPA Grand Challenge} \cite{wiki:darpachallenge} and the recent undertakings of most car manufacturers, such as BMW, Audi, Toyota, VW etc. \cite{forbes:timeline}

\newpar Certain ethical issues arise with autonomous vehicles, however. Since autonomous vehicles are programmed in advance to handle every possible scenario on the road, autonomous vehicles present ethical dilemmas that human drivers in non-autonomous vehicles do not, since an action has to be decided on \textit{before} the vehicle finds itself in an ethical dilemma. 

\newpar Human drivers are rarely blamed for acting according to their instinct in life-threatening vehicle collisions, even if this means that they inadvertently make decisions that have fatal consequences for other people. 

These decisions have to be programmed into autonomous vehicles in advance in order to make the "right" choice, should they occur. 
This raises interesting questions, such as; who should be blamed for an autonomous vehicle hitting, and possibly killing, one person over another? Should we even program autonomous vehicles to be able to select the "preferable" target of collision, or would it be better to not do any selection at all and make the vehicle choose some random behaviour in a critical situation?

The most interesting question, and therefore the one I will try to answer in this essay is; "Who is to blame for collisions or harm done to people or structures by an autonomous car in a critical situation, and can this blame even be placed on anyone?" by looking into what ethical issues this question presents, and what ethical theories that can help answer this question.

 

